### 2024-02-02 products/manuscript/Manuscript.qmd
first commit

---
title: "Class Exercise: Factors Influencing Abundance & Virus Isolation in Mosquitoes"
subtitle: "Which environmental, ecological, and temporal factors impact mosquito populations  "
author: Andrew Ruiz
date: "`r Sys.Date()`"
format:
  docx:
    toc: false
    number-sections: true
    highlight-style: github
bibliography: ../assets/dataanalysis-references.bib
csl: ../assets/apa.csl
editor: 
  markdown: 
    wrap: 72
---

```{r, echo=FALSE, message=FALSE}
# load a few R packages
library(here)
library(knitr)
library(readxl)

here::set_here()

```

# Summary/Abstract

Eastern equine encephalitis (EEE) is caused by an Alphavirus transmitted
to humans by the bite of an infected mosquito. Human infections are
rare, but serious. About 30-50% of symptomatic cases lead to death and
those who survive are left with life-long disability. The transmission
cycle for EEE includes two distinct cycles: an amplifying enzootic cycle
in which the virus is transmitted between the Culiseta melanura mosquito
and birds and an epizootic cycle in which bridge vector mosquitoes
transmit the EEE virus (EEEv) from birds to humans and other mammals.
Because C. melanura feeds almost exclusively on birds, it is not
considered a direct human threat. However, EEEv isolations in C.
melanura are an early warning sign that EEEv is circulating in the
ecosystem. Isolations in bridge vector mosquito species indicate
heightened transmission risk to humans.

This project will lay the foundation for my dissertation research. **The
goal for this phase of the project is to answer which environmental,
ecological, and temporal factors influence the transmission cycles for
EEE in southeastern Massachusetts.**

# Introduction

## General Background Information

Humans and other mammals are dead-end hosts for EEE. About 94% of human
EEE infections are asymptomatic. While symptomatic cases are rare, the
consequences can be severe. Mortality rates associated with symptomatic
infections range from 30 to 50%. Those who survive often experience
lifelong disability that includes debilitating neurological damage. On
average, there are 11 cases reported in the US annually. Most occur in
Massachusetts and Florida. Until recently, nearly all human EEE cases in
Massachusetts occurred in two counties: Bristol and Plymouth.

Since 2000, there have been 45 human cases recorded in Massachusetts,
resulting in 22 deaths. There are no human vaccines for EEE, and
treatment consists of palliative care only. Prevention measures include
personal behaviors to avoid mosquito bites and decreasing the mosquito
population through pesticide use and environmental modifications like
removing standing water. The Bristol County Mosquito Control Project
(BCMCP) coordinates mosquito surveillance and testing in the county from
June to October. BCMCP has used the same sentinel collection sites and
trapping methods for over 20 years. Once trapped, mosquitoes are sorted
by species, and vector species are submitted to the Massachusetts
Department of Health’s (MDPH) State Lab for PCR virus testing. Results
are available within 24 hours of submission. When rates of EEE mosquito
infections are above a defined threshold, MDPH notifies the local boards
of health and recommends preventive measures. Recommended measures
include outreach and education to increase personal protective practices
(Heymann, 2015). Additionally, MDPH may recommend that cities and towns
in the affected areas cancel evening outdoor events and discourage
outdoor activities when mosquito vectors are most active (dawn and dusk)
(Massachusetts Department of Public Health, n.d.). When infection rates
are high enough to indicate imminent human transmission, the
Commonwealth of Massachusetts will recommend and fund aerial pesticide
applications over the affected areas.

## Description of data and data source

*The cornerstone of this research will be the mosquito collection data
provided by a mosquito control project in SE Massachusetts. A formal
data-sharing agreement will be in place by January 2024. The dataset
encompasses nearly two decades of detailed mosquito surveillance. The
datasets include spatial identifiers. Mosquito collection data will be
integrated with other datasets using spatial or temporal attributes.
These datasets will include biological and ecological data,
environmental and spatial data, and demographic data, which are crucial
in understanding the transmission dynamics of EEE. Potential datasets
with their sources are outlined in the table below. Other datasets may
be included based on availability and need.*

| Data                         | Description                                                                                                                                           | Source                                |
|------------------|-------------------------------|-----------------------|
| Mosquito Surveillance Data   | Records of mosquito species counts, locations, and dates of collection. Data on the presence of EEE in mosquito populations is particularly valuable. | XXX County Mosquito Control           |
| Weather Data                 | Temperature, humidity, rainfall, and wind speed influence mosquito activity, population dynamics, and virus transmission.                             | NOAA                                  |
| Bird Population Data         | Since birds are a natural reservoir for EEE, information on bird populations and migration patterns could be relevant.                                | Crowd-sourced bird data               |
| Human Case Reports           | Data on confirmed human cases of EEE, including location, date of onset, and clinical outcomes.                                                       | MA Dept of Public Health publications |
| Veterinary Surveillance Data | Since horses and other mammals can also be affected by EEE, veterinary records could provide early warning signs of virus activity in an area.        | MA Dept of Public Health              |

### Currently, I have collected 2 datasets that are described below:

#### The first dataset is the mosquito collection dataset

```{r}
dp1 = here("data", "raw-data", "MOSQ_MADA.xlsx")

mos_data = read_excel(dp1)

head(mos_data)
summary(mos_data)

```

#### The second dataset is the weather data spanning the same time period as the mosquito collection data

```{r}
dp2 = here("data", "raw-data", "weather_airport.csv")
weather = read.csv(dp2, fileEncoding = "UTF-8")
head(weather)
summary(weather)
```

## Questions/Hypotheses to be addressed

\*This project will lay the foundation for my dissertation research.
**The goal for this phase of the project is to answer which
environmental, ecological, and temporal factors influence the
transmission cycles for EEE in southeastern Massachusetts.** While the
following questions will not be addressed in this phase, this project
will hopefully inform research that will answer the following questions:

1\. Can an AI-driven model that leverages historic mosquito
surveillance, ecological, and environmental data accurately quantify the
risk of human EEEv infection in southeastern Massachusetts?

2\. What is the potential of machine learning algorithms to identify
early warning signals for EEE outbreaks, enabling timely public health
interventions in Massachusetts to prevent human infections?

3\. How can machine learning models use real-time data on mosquito
activity, environmental factors, and weather to accurately predict when
the risk of EEE transmission to humans is low enough to lift evening
outdoor activity restrictions? \*

To cite other work (important everywhere, but likely happens first in
introduction), make sure your references are in the bibtex file
specified in the YAML header above (here
`dataanalysis_template_references.bib`) and have the right bibtex key.
Then you can include like this:

Examples of reproducible research projects can for instance be found in
[@mckay2020; @mckay2020a].

{{< pagebreak >}}

# Methods

*Specific methods will be determined when all data sources are
confirmed.*

## The process will involve creating a relational database in which all datasets are joined using spatial or temporal attributes. This will lead to a thorough examination of the datasets to understand the distributions, identify missing data and outliers, and uncover potential correlations between variables. The EDA will involve using visualizations, geographic information systems, and statistical tests to understand how biological, ecological, environmental, and demographic forces influence the EEE transmission cycles. The EDA will inform the preprocessing steps, feature engineering, and selection of appropriate AI models.

### Schematic of workflow

*This section will be updated when all data sources are confirmed*

## Data aquisition

With the exception of the mosquito collection data, all other datasets
are publicly available for use without restrictions.

## Data import and cleaning

*This step will be completed once all data sources are confirmed*

## Statistical analysis

This exercise will serve as the beginning of the exploratory data
analysis phase. One of the main objectives for of this phase is describe
the data, understand its structure, clean data, and identify
associations within the datasets.

{{< pagebreak >}}

# Results

## Exploratory/Descriptive analysis

*These will be generated in later phases of the exercise*

## Basic statistical analysis

## Full analysis

{{< pagebreak >}}

# Discussion

## Strengths and Limitations

## Conclusions

*Include citations in your Rmd file using bibtex, the list of references
will automatically be placed at the end*

This paper [@leek2015] discusses types of analyses.

These papers [@mckay2020; @mckay2020a] are good examples of papers
published using a fully reproducible setup similar to the one shown in
this template.

Note that this cited reference will show up at the end of the document,
the reference formatting is determined by the CSL file specified in the
YAML header. Many more style files for almost any journal [are
available](https://www.zotero.org/styles). You also specify the location
of your bibtex reference file in the YAML. You can call your reference
file anything you like, I just used the generic word `references.bib`
but giving it a more descriptive name is probably better.

{{< pagebreak >}}

# References