### 2024-03-12 products/manuscript/Manuscript.qmd
manuscript update
saving manuscript edits. began restructuring the sections.

---
title: "DRAFT: Factors Influencing Abundance & Virus Isolation in Mosquitoes"
subtitle: "Which environmental, ecological, and temporal factors impact mosquito populations  "
author: Andrew Ruiz
date: "`r Sys.Date()`"
format:
  docx:
    toc: false
    number-sections: true
    highlight-style: github
editor: 
  markdown: 
    wrap: 72
---

```{r, echo=FALSE, message=FALSE}
# load a few R packages
library(here)
library(knitr)
library(readxl)
library(dplyr)
library(kableExtra)
library(dplyr)
library(readr)

here::set_here()

```

# Summary/Abstract

Eastern equine encephalitis (EEE) is caused by an Alphavirus transmitted
to humans by the bite of an infected mosquito. Human infections are
rare, but serious. About 30-50% of symptomatic cases lead to death and
those who survive are left with life-long disability. The transmission
cycle for EEE includes two distinct cycles: an amplifying enzootic cycle
in which the virus is transmitted between the Culiseta melanura mosquito
and birds and an epizootic cycle in which bridge vector mosquitoes
transmit the EEE virus (EEEv) from birds to humans and other mammals.
Because C. melanura feeds almost exclusively on birds, it is not
considered a direct human threat. However, EEEv isolations in C.
melanura are an early warning sign that EEEv is circulating in the
ecosystem. Isolations in bridge vector mosquito species indicate
heightened transmission risk to humans.

# Problem statement

Which environmental,ecological, and temporal factors influence the
transmission cycles for EEE in southeastern Massachusetts and which
models best describe the relationship between these factors.

This project will lay the foundation for dissertation research that will
begin in June, 2024. The research will focus on three

# Introduction

## Background

Humans and other mammals are dead-end hosts for EEE. About 94% of human
EEE infections are asymptomatic. While symptomatic cases are rare, the
consequences can be severe. Mortality rates associated with symptomatic
infections range from 30 to 50%. Those who survive often experience
lifelong disability that includes debilitating neurological damage. On
average, there are 11 cases reported in the US annually. Most occur in
Massachusetts and Florida. Until recently, nearly all human EEE cases in
Massachusetts occurred in two counties: Bristol and Plymouth.

Since 2000, there have been 45 human cases recorded in Massachusetts,
resulting in 22 deaths. There are no human vaccines for EEE, and
treatment consists of palliative care only. Prevention measures include
personal behaviors to avoid mosquito bites and decreasing the mosquito
population through pesticide use and environmental modifications like
removing standing water.

The Bristol County Mosquito Control Project (BCMCP) coordinates mosquito
surveillance and testing in the county from June to October. BCMCP has
used the same sentinel collection sites and trapping methods for over 20
years. Once trapped, mosquitoes are sorted by species, and vector
species are submitted to the Massachusetts Department of Health’s (MDPH)
State Lab for PCR virus testing. Results are available within 24 hours
of submission. When rates of EEE mosquito infections are above a defined
threshold, MDPH notifies the local boards of health and recommends
preventive measures. Recommended measures include outreach and education
to increase personal protective practices . Additionally, MDPH may
recommend that cities and towns in the affected areas cancel evening
outdoor events and discourage outdoor activities when mosquito vectors
are most active (dawn and dusk). When infection rates are high enough to
indicate imminent human transmission, the Commonwealth of Massachusetts
will recommend and fund aerial pesticide applications over the affected
areas.

# Methods

## Study area

Bristol County, Massachusetts covers an area of 691 square miles with a
population density of 962 people per square mile. It borders Rhode
Island to the west, Plymouth County, MA to the east, and Norfolk County,
MA to the north.The county is situated on Buzzard's Bay to the south.

Atlantic White Cedar (AWC) swamps span over 4,300 acres within
Massachusetts, primarily concentrated in Plymouth and Bristol counties.
These swamps serve as critical breeding habitats for Culiseta melanura
(MEL), which is recognized as the primary enzootic vector for Eastern
Equine Encephalitis (EEE). AWC trees have large roots that form "crypts"
that hold water throughout the year. These crypts serve as a main
habitat for MEL larval growth.

Between 2003 and 2022 Plymouth and Bristol counties were the epicenters
for EEE transmission in Massachusetts, accounting for 24 out of the
state's 41 human EEE cases.The Bristol County Mosquito Control Project
(BCMCP) conducts mosquito surveillance and abatement activities for the
county and often coordinates activities with Plymouth County Mosquito
Control Project.

## Data collection and processing

The cornerstone of this project is mosquito surveillance data. This
dataset includes mosquito counts by species for all mosquito traps set
in Bristol County, MA between 2007 to 2023. In addition to mosquito
counts by species, the dataset includes the town or city in which they
were collected, the collection data, and whether the mosquitoes were
submitted for virus testing.

While the surveillance data provide insight into mosquito population
estimates and changes in species composition, it is missing one key
feature. The virus isolation results are not included in this dataset.
However, virus isolations in mosquitoes are published on the MDPH
website by town/city and collection date. With this information, one can
identify the trap event contained in the dataset which allow for
calculations of key measure, such as minimum infection rate (MIR).

The surveillance dataset is structured to conform with state and federal
reporting requirements. As a result, the format of the data must be
transformed to in order to make it ready for analysis. Once the data
structure is corrected, it can be joined to other relevant datasets
using spatial or temporal attributes.

This dataset will be integrated with other datasets using spatial or
temporal attributes. These datasets will include biological and
ecological data, environmental and spatial data, and demographic data,
which are crucial in understanding the transmission dynamics of EEE.
Potential datasets with their sources are outlined in the table below.
Other datasets may be included based on availability and need.

| Data                         | Description                                                                                                                                           | Source                                |
|-------------------|----------------------------------|-------------------|
| Mosquito Surveillance Data   | Records of mosquito species counts, locations, and dates of collection. Data on the presence of EEE in mosquito populations is particularly valuable. | XXX County Mosquito Control           |
| Weather Data                 | Temperature, humidity, rainfall, and wind speed influence mosquito activity, population dynamics, and virus transmission.                             | NOAA                                  |
| Bird Population Data         | Since birds are a natural reservoir for EEE, information on bird populations and migration patterns could be relevant.                                | eBird                                 |
| Human Case Reports           | Data on confirmed human cases of EEE, including location, date of onset, and clinical outcomes.                                                       | MA Dept of Public Health publications |
| Veterinary Surveillance Data | Since horses and other mammals can also be affected by EEE, veterinary records could provide early warning signs of virus activity in an area.        | MA Dept of Public Health              |

## Questions/Hypotheses to be addressed

\*This project will lay the foundation for my dissertation research.
**The goal for this phase of the project is to answer which
environmental, ecological, and temporal factors influence the
transmission cycles for EEE in southeastern Massachusetts.**

While the following questions will not be addressed in this phase, this
project will hopefully inform research that will answer the following
questions:

1\. Can an AI-driven model that leverages historic mosquito
surveillance, ecological, and environmental data accurately quantify the
risk of human EEEv infection in southeastern Massachusetts?

2\. What is the potential of machine learning algorithms to identify
early warning signals for EEE outbreaks, enabling timely public health
interventions in Massachusetts to prevent human infections?

3\. How can machine learning models use real-time data on mosquito
activity, environmental factors, and weather to accurately predict when
the risk of EEE transmission to humans is low enough to lift evening
outdoor activity restrictions? \*

{{< pagebreak >}}

# Methods

*Specific methods will be determined when all data sources are
confirmed.*

## The process will involve creating a relational database in which all datasets are joined using spatial or temporal attributes. This will lead to a thorough examination of the datasets to understand the distributions, identify missing data and outliers, and uncover potential correlations between variables. The EDA will involve using visualizations, geographic information systems, and statistical tests to understand how biological, ecological, environmental, and demographic forces influence the EEE transmission cycles. The EDA will inform the preprocessing steps, feature engineering, and selection of appropriate AI models.

### Schematic of workflow

*This section will be updated before analysis begins*

## Data aquisition

With the exception of the mosquito collection data, all other datasets
are publicly available for use without restrictions.

## Data import and cleaning

Datasets will be loaded into R in a Quarto document to document and
describe cleaning methods. There are two datasets that will require more
extensive cleaning: the mosquito trap data and the crowd-sourced bird
data. The process is outlined in detail in the MADAproject_pt2.qmd file.
Less significant cleaning and data management steps not included in this
manuscript are also outlined in that file.

The mosquito dataset spans 2007 to 2023. However, the virus and case
data only covers the years 2014 to 2020. Because mosquito populations
can be influences by processes of the previous year. We will set the
analytical horizon will span 2013 to 2021.

Additionally, the Massachusetts Department of Public Health requires
that all mosquito control district submit their colelction data in a
specific format. This format divides each trap event into separate rows
based on mosquito species. However, this format is not useful for
certain calculations where a zero count for a species is not explicitly
recorded. For example, even when sorted by species to calculate the
average count per trap, the average could be inflated since the records
will not include trap events with a zero count of that species. In order
to correct this, the data was pivoted to create a column for every
unique "species code" then the "pool size" for each species code is
summed for each trap event. A trap event is when "town", "date of
collection", "trap type" are all the same.. For species codes not
included in the orginal table for a trap event, the field will be
assigned a zero value. The resulting table will have one row for every
trap event.

The bird observation dataset was obtained from Cornell University's
eBird program. The original dataset included all reported bird
observations for Southeast Massachusetts and was over 600 MB of data.
While all birds are potential amplifying hosts for EEE and West Nile
Virus (WNV), birds from the Genus *Corvus* are of particular concern.
The bird dataset processed to include records of *Corvus spp* sightings
after 2014. The final dataset included 8 of the 50 variables, including
location data.

## Statistical analysis

This exercise will serve as the beginning of the exploratory data
analysis phase. One of the main objectives for of this phase is describe
the data, understand its structure, clean data, and identify
associations within the datasets.

# Results

## Exploratory/Descriptive analysis

The first phase of the exploration is complete. The next phase will
include statistical tests to determine if there are possible
associations between the environmental variables and the mosquito
abundance as measured by trap counts or EEE virus activity.

It is important to know where the mosquitoes infected with EEE are
captured. Below is a table showing the virus isolations for EEE in
trapped mosquitoes by county between 2014 and 2020.

![](images/eee_isolation_county.png){fig-align="center"}

The geographic focus for this study is Bristol County. One factor to
consider is if early virus isolations are associated with cases in
humans and other mammals. The table below shows the date of the first
EEE virus isolation in mosquitoes and the number of human of other
animal cases by year.

![](images/eee_isolation_cases-03.png){fig-align="center"}

Unfortunately, the time horizon is limited by data availability.
Additionally, EEE human infections are rare, so traditional statistical
methods may be inadequate to suggest any relationships. What is clear is
that 2019 was an exceptional year for EEE.

*Culiseta melanura* (MEL) plays an important role in EEE virus
amplification in the enzootic cycle. Understanding the population
dynamics might provide more insight into what is driving the 2019 spike.

The graph below shows the proportion of MEL among all mosquito vector
species trapped in Bristol County, MA.

![](images/mel_proportion.png){fig-align="center"}

From this we see a large proportion of MEL in the traps for 2019. In
fact, they comprise more than half of all vector species for that year.
Another interesting aspect is the high proportion in 2018. MEL
overwinters as larvae and as long as there is sufficient standing water
to last the winter, MEL hatched in the previous fall would survive to
emerge the next year. This suggests it might be important to look at
conditions in the previous year or season for a better understanding.

Since water plays a significant role in the mosquito life cycle, let's
look at the annual precipitation.

![](images/annual_precip-02.png){fig-align="center"}

The graph indicates that the annual precipitation total for 2018 is
higher than the other years. This suggest that there was enough water to
support large MEL larval populations though out the winter.

## Basic statistical analysis

*These will be generated in later phases of the exercise*

## Full analysis

*These will be generated in later phases of the exercise* \# Discussion
*this section will be completed when analysis is done*

## Strengths and Limitations

*this section will be updated after analysis is complete*

While the reliability of the mosquito trap data is high, the trap data
cannot be directly associated with the virus isolation data. The
information this scale is not released by the State. Spatial scale is
also an issue with the human and veterinary cases data. Human case data
is released only on a county scale to ensure patient privacy. Some
veterinary case data is available on the town-level, but most remains
only on the county-level.

These limitations are barriers to accurate spatial analysis.

## Conclusions *this section will be completed when analysis is done*

# References

References will be added in part 3 of the project